---
title: "Quantile correlation"
author: "Jiaxi Li, Marlon Do Couto"
date: "2/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Rolling Quantile Correlation Graphs

* Purpose: Explore and visualize quantile correlations among key financial assets

* Method: Calculate the rolling quantile correlation of various ETFs by first using their daily returns and rolling it through a three year window (using 252 day convention for each year). The quantiles chosen are **0.05, 0.25, 0.50, 0.75 and 0.95** which were calculated using the QCSIS's qc() function^1. 

* Data: Daily adjusted returns for the twelve main ETFs since their inception. 

* Authors: 

  + [Marlon Do Couto (December 2020 - Present)](https://www.linkedin.com/in/marlondocouto/)
  
  + [Scarlett Meng (December 2020 - Present)](https://www.linkedin.com/in/scarlettmeng/)

  + [Jiaxi Li (December 2020 - Present)](https://www.linkedin.com/in/jiaxi-li-b0389111a/)

[1] https://rdrr.io/cran/QCSIS/man/QCSIS-package.html 


```{r, message=FALSE, warning=FALSE}
library(tidyverse) # contains ggplot2 tidyr dplyr
library(lubridate)
library(tidyquant)
library(QCSIS) # for quantile correlation
library(tibbletime) # rollapply fuction
library(ggthemes)

Main_ETFs <- c("IVV","IJH","IJR","AGG","TLT","IEFA","IEMG","IAU","GSG","FXE","FXB","FXY")
Core_prices <- tq_get(Main_ETFs, get = "stock.prices")
returns <- Core_prices %>%
  group_by(symbol)%>%
  tq_transmute(select = adjusted,
               mutate_fun = periodReturn,
               period="daily",
               col_rename = "ret")%>%na.omit()

rolling_quant_corr<-function(returns){
  Combined <- returns %>%
    rename(symbol1 = symbol,ret1 = ret)%>%
    inner_join(returns %>%
                 rename(symbol2 = symbol,
                      ret2 = ret))
  # Apply rollify function and qc function to create the rolling qc functions for 3 year trading window:
  roll_qc_05 <- rollify(~ qc(.x, .y, 0.05)$rho, window = 252*3)
  roll_qc_25 <- rollify(~ qc(.x, .y, 0.25)$rho, window = 252*3)
  roll_qc_50 <- rollify(~ qc(.x, .y, 0.5)$rho, window = 252*3)
  roll_qc_75 <- rollify(~ qc(.x, .y, 0.75)$rho, window = 252*3)
  roll_qc_95 <- rollify(~ qc(.x, .y, 0.95)$rho, window = 252*3)

  Corr.Matrix <- Combined %>%
    ungroup() %>%
    group_by(symbol1, symbol2) %>%
    mutate(`0.05` = roll_qc_05(ret2, ret1),
           `0.25` = roll_qc_25(ret2, ret1),
           `0.50` = roll_qc_50(ret2, ret1),
           `0.75` = roll_qc_75(ret2, ret1),
           `0.95` = roll_qc_95(ret2, ret1))%>%
    na.omit()%>%
    pivot_longer(`0.05`:`0.95`, names_to = "Quantile", values_to = "QC")
  
  return(Corr.Matrix)
}

Corr.Matrix<-rolling_quant_corr(returns)

```


```{r}
plot_roll_corr<-function(data){
  graph.list<-list()
  count<-0

  for (i in unique(data$symbol1)){
    count<-count+1
    graph.list[[count]]<- Corr.Matrix%>%filter(symbol1==i)%>%filter(symbol2!=i)%>%
      mutate(symbol2=factor(unique(symbol2)))%>%ggplot(aes(x=date,y=QC,color=factor(Quantile)))+
      geom_line()+theme_economist()+
      guides(color=guide_legend(override.aes = list(size=2)))+
      scale_color_manual(values = c("#e6194B","#ffe119","#000075","#3cb44b","#f58231"))+
      facet_wrap(~symbol2)+labs(title=paste("Correlations of ",i),x=" ",y=" ",color="Quantile")
    }
  return(graph.list)
}

graphs<-plot_roll_corr(Corr.Matrix)

print(graphs[1:12])
```
```{r}
Combined<-returns%>%rename(symbol1=symbol, ret1=ret)%>%
  inner_join(returns%>%rename(symbol2=symbol, ret2= ret))

rolling<-rollify(~cor(.x,.y),window = 252*3)

Normal.Corr<-Combined%>%filter(symbol1=="IVV")%>%filter(symbol2=="IJR")%>%
  mutate(cor=rolling(ret2,ret1))%>%na.omit()
  

corr2<- Normal.Corr%>%ggplot(aes(x=date,y=cor))+
  geom_line()+theme_economist()+
  guides(color=guide_legend(override.aes = list(size=2)))+
  labs(title="Traditional Correlations of IVV and IJR",x=" ",y=" ")

corr2
ggsave("ivv_ijr_normal.png",corr2)
```

