---
title: "Technology vs Tourism Stocks Pre and Post Pandemic"
author: "Marlon Do Couto"
date: "1/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this case is to investigate the divergence between the technology and travel/leisure stocks prior to Covid-19 and after Covid-19. To do the analysis, I first pick ETFs that track both areas:

**Travel and leisure**: I will divide  area into two main areas: *airlines* and *hotels*. To track the airline industry, I will use *JETS*; This ETF is totally invested in airlines and aviation with and 80/20 domestic and international allocation. To track hotels, I will also create a simple index of hotels/entertainment equally allocation with:
- Marriott International (*MAR*)
- Hilton Hotels (*HLT*)
- Hyatt Hotels (*H*)
- Choice Hotels (*CHH*)
- Wyndham Hotels & Resorts (*WH*)

**Technology**: For technology companies, we will use *QQQ* which has large holdings in the FANGs and other technology companies such as Tesla, NVIDIA, etc.

First step, we will get our data using tidyquant. We will use data from 2019-01-01 so that we have observations pre and post pandemic. 

```{r}
library(tidyquant)
library(tidyverse)
library(timetk)

date<-"2019-01-01"

airlines.data<-tq_get("JETS",from=date,get="stock.prices")
hotels<-c('MAR','HLT','H','CHH','WH')
hotels.data<-tq_get(hotels,from=date,get="stock.prices")
technology.data<-tq_get("QQQ",from=date,get="stock.prices")

head(airlines.data)
head(hotels.data)
head(technology.data)
```

Now, let's create an equally-weighted portfolio of the hotel stocks:
```{r}
weights<-c(0.2,0.2,0.2,0.2,0.2)

hotel.returns<-hotels.data%>%group_by(symbol)%>%tq_transmute(adjusted,periodReturn,period="daily",                                              col_rename="ret")
hotel.index<-tq_portfolio(data=hotel.returns,assets_col = symbol,returns_col = ret,
                          weights = weights,col_rename ="ret", wealth.index=FALSE) %>%
  mutate(ret=if_else(row_number()==1,0,ret))%>%
  mutate(cum.ret=cumprod(1+ret)-1)
hotel.index$symbol<-"Hotels"
hotel.index<-hotel.index[,-2]
head(hotel.index)


```
Now, we calculate the returns for *JETS* and *QQQ* as well:

```{r}
tech.returns<-technology.data%>%tq_transmute(adjusted,periodReturn,period="daily",col_rename = "ret")%>%
  mutate(ret=if_else(row_number()==1,0,ret))%>%mutate(cum.ret=cumprod(1+ret)-1)
tech.returns$symbol<-"QQQ"
tech.returns<-tech.returns[,-2]

jets.returns<-airlines.data%>%tq_transmute(adjusted,periodReturn,period="daily",col_rename = "ret")%>%
  mutate(ret=if_else(row_number()==1,0,ret))%>%mutate(cum.ret=cumprod(1+ret)-1)
jets.returns$symbol<-"JETS"
jets.returns<-jets.returns[,-2]

head(tech.returns)
head(jets.returns)
```

Now, we put together the returns into a single table:

```{r}
all.returns<-rbind(hotel.index,tech.returns,jets.returns)
head(all.returns)
```



Now, let's graph the returns to see the changes:

```{r}

all.returns%>%ggplot(aes(x=date,y=cum.ret,color=symbol))+geom_line()+
  theme_classic()+scale_y_continuous(breaks=seq(-0.5,1.2,0.1), labels = scales::percent)+
  coord_cartesian(clip="off")+labs(x="Year-Month",y="Returns",title="QQQ, JETS, and Hotels Returns since Jan 01, 2019")
```
Now, using an animated chart with the gganimate package.

```{r}
library(gganimate)
library(gifski)

animatedplot<-all.returns%>%ggplot(aes(x=date,y=cum.ret,group=symbol,color=symbol))+
  geom_line()+
  geom_segment(aes(xend=ymd("2021-01-29"),yend=cum.ret),linetype=2)+
  geom_text(aes(x=ymd("2021-01-29"),label=symbol),hjust=0)+
  theme_classic()+theme(legend.position = "none")+
  scale_y_continuous(breaks=seq(-0.5,1.2,0.1), labels = scales::percent)+
  transition_reveal(date)+
  labs(x="Date",y="Returns",title="QQQ, JETS and Hotels Returns since Jan 01, 2019")+ease_aes('linear')

animate(animatedplot,duration = 5,fps=20,renderer = gifski_renderer())
anim_save("tech_tourism.gif")
```

